The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-08-24 article.tex
Placeholder section for hash codes
Misnomer to call it MDL molfile, correct casing (no capital m) in my section.

  The perception from coordinates and wedge assignment algorithms are fundamental for conversion
  between formats that store stereochemistry implicitly based on coordinates (e.g. Molfile, CML) and 
  explicitly (e.g. SMILES, CML, InChI). Perception from 2D coordinates can optionally identify
  perspective projects, specifically: Fischer, Haworth, and Chair projections. With the perception of 
  perspective projections enabled, database entries currently consider distinct can be merged
  (Figure~3\ref{fig:stereoprojections}). The perception is based on an algorithm briefly 
  described by \cite{Karapetyan2015}.

@@

  \subsubsection*{MDL molfile improvements}

The MDL molfile formats are still very popular, despite it being a proprietary
standard, but taking advantage of the detailed specification still floating
around on the world wide web and can be found by googling for
ctfile.pdf~\cite{ctfile.pdf}. The CDK has various related readers, reflecting
the various formats specified in this document, as well as a more historical
reader for the pre-V2000 formats. When this reader was written, MDL molfiles
in this pre-V2000 format were still around, but the authors are unaware of
existing use of this format. At the same time, V3000 does not have seen a
significant adoption either. Furthermore, several formats exist that wrap
molfiles in collections, such as structure data files (SD files), reaction files,
and reaction collections. Table XXX shows an overview of the various readers
for (partially) supported formats.
% note, I pie chart would be cool, showing the various relative amounts of the
% various formats.

@@

@@

Since the last CDK publication in 2006, the library has been improved in many
aspects. The functionality has seen many additions and rewrites, making the
library more functional and a lot more faster then before. Updates on the common
SMILES and MDL molfile formats and the improved structure diagram generation are
very visible and benefit many of the tools using the CDK.
Furthermore, the stability of the development model has significantly improved,
providing greater stability of the library over time. The project adopted peer
review and quality control.

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lp{3cm}p{3cm}l}
  \textbf{Module}            & \textbf{Description}  & \textbf{Major Changes} & \textbf{Dependencies} \\ \hline
  interfaces                 & Interfaces for the data models. & & Vecmath 1.5.2 \\ \hline
  core                       & Core functionality.             & & Google Guava 17.0 \\ \hline % FIXME: how to add https://github.com/google/guava ?
  standard                   & Common functionality.           & & \\ \hline
  render                     & Graphical rendering.            & Redesigned to make it more modular and support multiple widget toolkits, like AWT and SWT. & \\ \hline
  isomorphism                & Isomorphism and substructure searching. & & \\ \hline
  atomtype                   & Various non-core atom type schemes.     & Unified approach where atom typing is separated from other algorithms. & \\ \hline
  ioformats                  & Definitions of (chemical) input/output formats. & & \\ \hline
  io                         & Readers and writers for input/output formats.  & The MDL molfile reader has been rewritten and supports in atom types defind in the specification. & XPP3 1.1.4c \\ \hline
  iordf                      & Stores data models as in the Resource Description Framework serialization foramts. & New. & Jena 2.7.4 \\ \hline
  inchi                      & IUPAC International Chemical Identifier support. & & JNI-InChI 0.8~\cite{Spjuth2013}  \\ \hline
  libiocml                   & Writer for the Chemical Markup Language format. & & XOM 1.2.5, CMLXOM 3.1~\cite{Murray-Rust2011} \\ \hline
  sdg                        & Structure diagram generation.  & Much improved overlap resolution. & \\ \hline
  smiles                     & Reading and writing in the SMILES format. & SMILES support performance and converage is greatly improved. & Beam 0.9.1~\cite{Beam} \\ \hline
  smarts                     & Substructure searching with the SMARTS format. & & Beam 0.9.1~\cite{Beam} \\ \hline
  formula                    & Chemical formula support. & New. & \\ \hline
  fingerprint                & Calculate fingerprints. & Many new fingerprint types (see text). & Apache Commons Math 3.1.1 \\ \hline
  qsar and qsarmolecular                       & Molecular descriptors.  & & XOM 1.2.5, JAMA 1.0.3~\cite{Hicklin2012} \\ \hline
  signatures                 & Calculation of molecular and atomic signatures. & & Signatures 1.1 \\ \hline % FIXME: ask Gilleain for the reference
    \end{tabular}
    \end{minipage}

>>>>

  The perception from coordinates and wedge assignment algorithms are fundamental for conversion
  between formats that store stereochemistry implicitly based on coordinates (e.g. molfile, CML) and 
  explicitly (e.g. SMILES, CML, InChI). Perception from 2D coordinates can optionally identify
  perspective projects, specifically: Fischer, Haworth, and Chair projections. With the perception of 
  perspective projections enabled, database entries currently consider distinct can be merged
  (Figure~3\ref{fig:stereoprojections}). The perception is based on an algorithm briefly 
  described by \cite{Karapetyan2015}.

@@

\subsubsection*{molfile improvements}

The molfile formats are still very popular, despite it being a proprietary
standard, but taking advantage of the detailed specification still floating
around on the world wide web and can be found by googling for
ctfile.pdf~\cite{ctfile.pdf}. The CDK has various related readers, reflecting
the various formats specified in this document, as well as a more historical
reader for the pre-V2000 formats. When this reader was written, molfiles
in this pre-V2000 format were still around, but the authors are unaware of
existing use of this format. At the same time, V3000 does not have seen a
significant adoption either. Furthermore, several formats exist that wrap
molfiles in collections, such as structure data files (SD files), reaction files,
and reaction collections. Table XXX shows an overview of the various readers
for (partially) supported formats.
% note, I pie chart would be cool, showing the various relative amounts of the
% various formats.

@@

\subsubsection*{Hash Codes}

%% JWM todo

@@

Since the last CDK publication in 2006, the library has been improved in many
aspects. The functionality has seen many additions and rewrites, making the
library more functional and a lot more faster then before. Updates on the common
SMILES and molfile formats and the improved structure diagram generation are
very visible and benefit many of the tools using the CDK.
Furthermore, the stability of the development model has significantly improved,
providing greater stability of the library over time. The project adopted peer
review and quality control.

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lp{3cm}p{3cm}l}
  \textbf{Module}            & \textbf{Description}  & \textbf{Major Changes} & \textbf{Dependencies} \\ \hline
  interfaces                 & Interfaces for the data models. & & Vecmath 1.5.2 \\ \hline
  core                       & Core functionality.             & & Google Guava 17.0 \\ \hline % FIXME: how to add https://github.com/google/guava ?
  standard                   & Common functionality.           & & \\ \hline
  render                     & Graphical rendering.            & Redesigned to make it more modular and support multiple widget toolkits, like AWT and SWT. & \\ \hline
  isomorphism                & Isomorphism and substructure searching. & & \\ \hline
  atomtype                   & Various non-core atom type schemes.     & Unified approach where atom typing is separated from other algorithms. & \\ \hline
  ioformats                  & Definitions of (chemical) input/output formats. & & \\ \hline
  io                         & Readers and writers for input/output formats.  & The molfile reader has been rewritten and supports in atom types defind in the specification. & XPP3 1.1.4c \\ \hline
  iordf                      & Stores data models as in the Resource Description Framework serialization foramts. & New. & Jena 2.7.4 \\ \hline
  inchi                      & IUPAC International Chemical Identifier support. & & JNI-InChI 0.8~\cite{Spjuth2013}  \\ \hline
  libiocml                   & Writer for the Chemical Markup Language format. & & XOM 1.2.5, CMLXOM 3.1~\cite{Murray-Rust2011} \\ \hline
  sdg                        & Structure diagram generation.  & Much improved overlap resolution. & \\ \hline
  smiles                     & Reading and writing in the SMILES format. & SMILES support performance and converage is greatly improved. & Beam 0.9.1~\cite{Beam} \\ \hline
  smarts                     & Substructure searching with the SMARTS format. & & Beam 0.9.1~\cite{Beam} \\ \hline
  formula                    & Chemical formula support. & New. & \\ \hline
  fingerprint                & Calculate fingerprints. & Many new fingerprint types (see text). & Apache Commons Math 3.1.1 \\ \hline
  qsar and qsarmolecular                       & Molecular descriptors.  & & XOM 1.2.5, JAMA 1.0.3~\cite{Hicklin2012} \\ \hline
  signatures                 & Calculation of molecular and atomic signatures. & & Signatures 1.1 \\ \hline % FIXME: ask Gilleain for the reference
    \end{tabular}
    \end{minipage}
